Wheels of Misfortune: Exploring Auto Theft Trends in Ottawa

### 2024-04-22 paper/paper.qmd
updated graphs

```{r}
#| label: theft-year
#| fig-cap: "Average Auto Theft Crime Reported in ottawa from 2018-2023"
#| echo: false
#| warning: false
#| message: false

auto_Data |>
  ggplot(aes(x = theft_year)) +
  geom_histogram(stat = "count")
```
```{r}
#| label: theft-brand
#| fig-cap: "Number of Cars Stolen by Car Brand in 2018"
#| echo: false
#| warning: false
#| message: false

brand_theft_2018 <- auto_Data |>
  filter(theft_year == "2018")

brand_theft_2018 |>
  ggplot(aes(x = vehicle_make, fill = vehicle_make)) +
  geom_histogram(stat = "count")

@@

```{r}
#| echo: false
#| warning: false
#| message: false

# creating map from section 5.4.1 in the textbook

# Define the bounding box for Ottawa
bbox <- c(left = -75.9, bottom = 45.3, right = -75.5, top = 45.5)

# Get the map of Ottawa using get_stadiamap function
ottawa_stadia_map <- get_stadiamap(bbox, zoom = 11, maptype = "stamen_toner_lite")

# Plot the map using ggmap
ggmap(ottawa_stadia_map) # map prints

# ---- not working -----------
#filtering dataset to get auto theft reports per neighborhood in 2018

#neighborhood_theft_2018 <- auto_Data |>
#  filter(theft_year == "2018") |>
#  select(neighbourhood, x, y)

#print(neighborhood_theft_2018)

# map does not print with elements plotted

#ggmap(ottawa_stadia_map, extent = "normal", maprange = FALSE) +
#  geom_point(data = neighborhood_theft_2018,
#             aes(x = 45.3, y = 45.5, colour = neighbourhood),
#             alpha = 0.7) +
#  scale_color_brewer(name = "2019 Division", palette = "Set1") +
#  labs(x = "Longitude",
#       y = "Latitude") +
#  theme_minimal() +
#  theme(panel.grid.major = element_blank(),
#        panel.grid.minor = element_blank())

@@

# Model details {#sec-model-details}

>>>>

```{r}
#| label: averageTheftYear
#| fig-cap: "Average Auto Theft Crime Reported in ottawa from 2018-2023"
#| echo: false
#| warning: false
#| message: false

auto_Data |>
  ggplot(aes(x = theft_year)) + 
  labs(x = "Year", y = "Count") + theme_minimal() +
  geom_histogram(stat = "count")

```
```{r}
#| label: theft-brands
#| fig-cap: "Number of Cars Stolen by Car Brand in 2018"
#| fig-subcap: ["Illnesses","Visits to the doctor"]
#| layout-ncol: 3
#| echo: false
#| warning: false
#| message: false

# Filter for Number of Cars Stolen by Brand in 2018 
brand_theft_2018 <- auto_Data |>
  filter(theft_year == "2018")

# filter for Top 10 Brands 
brand_theft_2018_table <- brand_theft_2018 |>
                          count(vehicle_make)

brands2018_top10 <- brand_theft_2018_table |>
  filter(n > 26, n != 95) |> 
  mutate(vehicle_make = recode(vehicle_make,
                             "CHEV" = "Chevrolet",
                             "DODG" = "Dodge",
                             "FORD" = "Ford",
                             "GMC" = "GMC",
                             "HOND" = "Honda",
                             "HYUN" = "Hyundai",
                             "LEXS" = "Lexus",
                             "MAZD" = "Mazda",
                             "NISS" = "Nissan",
                             "TOYT" = "Toyota"
                             ))
brands2018_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2018")

# Filter for Number of Cars Stolen by Brand in 2019 
brand_theft_2019 <- auto_Data |>
  filter(theft_year == "2019")

# filter for Top 10 Brands 
brandtheft2019_top10 <- brand_theft_2019 |>
                          count(vehicle_make) |>
                          filter(n > 25, n != 118) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2019_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2019")

# Filter for Number of Cars Stolen by Brand in 2020
brand_theft_2020 <- auto_Data |>
  filter(theft_year == "2020")

# filter for Top 10 Brands 
brandtheft2020_top10 <- brand_theft_2020 |>
                          count(vehicle_make) |>
                          filter(n > 21, n != 95) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2020_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2020")

# Filter for Number of Cars Stolen by Brand in 2021 
brand_theft_2021 <- auto_Data |>
  filter(theft_year == "2021")

# filter for Top 10 Brands 
brandtheft2021_top10 <- brand_theft_2021 |>
                          count(vehicle_make) |>
                          filter(n > 21, n != 115) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2021_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2021")

# Filter for Number of Cars Stolen by Brand in 2022
brand_theft_2022 <- auto_Data |>
  filter(theft_year == "2022")

# filter for Top 10 Brands 
brandtheft2022_top10 <- brand_theft_2022 |>
                          count(vehicle_make) |>
                          filter(n > 34, n != 124) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2022_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022")

# Filter for Number of Cars Stolen by Brand in 2023 
brand_theft_2023 <- auto_Data |>
  filter(theft_year == "2023")

# filter for Top 10 Brands 
brandtheft2023_top10 <- brand_theft_2023 |>
                          count(vehicle_make) |>
                          filter(n > 16, n != 36) |> 
                          mutate(vehicle_make = recode(vehicle_make,
                                                       "ACUR" = "Acura",
                                                        "CHEV" = "Chevrolet",
                                                        "DODG" = "Dodge",
                                                        "FORD" = "Ford",
                                                        "GMC" = "GMC",
                                                        "HOND" = "Honda",
                                                        "HYUN" = "Hyundai",
                                                        "LEXS" = "Lexus",
                                                        "MERZ" = "Mercedes-Benz ",
                                                        "MAZD" = "Mazda",
                                                        "NISS" = "Nissan",
                                                        "TOYT" = "Toyota"
                                                     ))
brandtheft2023_top10 |>
  ggplot(aes(x = vehicle_make, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2023")

@@

```{r}
#| label: theft-recovery
#| fig-cap: "Recovery Rates from 2018 - 2023"
#| echo: false
#| warning: false
#| message: false

theft_recovery <- auto_Data |>
  select(theft_year, recovered)

theft_recovery |>
  ggplot(mapping = aes(x = theft_year, fill = recovered)) +
  geom_bar(position = "dodge2") +
  labs(x = "Age group", y = "Number of observations", fill = "Vote") +
  theme(legend.position = "bottom")
```
```{r}
#| label: theft-neighborhoods
#| fig-cap: "Recovery Rates from 2018 - 2023"
#| echo: false
#| warning: false
#| message: false

theft_recovery <- auto_Data |>
  select(neighbourhood) |> 
  count(neighbourhood) |> 
  filter(n > 180)

theft_recovery |>
  ggplot(aes(x = neighbourhood, y = n, fill = n)) +  # Use n as the fill aesthetic
  geom_bar(stat = "identity") +
  theme_minimal() + theme(legend.position = "bottom") + 
  labs(x = "Vehicle Brand", y = "Count", fill = "Count", title = "Top 10 Brands with Most Auto Thefts in 2022") 
```

@@
